UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-07-13 ms/main.tex
small edit

With these caveats in mind, a priori data visualization plays a central role in quality control, hypothesis generation, and confounder identification for a wide range of genomic applications. Nonlinear approaches, despite their limitations, become increasingly useful as the size of datasets increases. UMAP, in particular, reveals a wide range of features that would not be apparent using linear maps. Given its ease of use, breadth of results, and low computational cost, we propose that UMAP should become a default companion to PCA in large genomic cohorts.

>>>>

With these caveats in mind, a priori data visualization plays a central role in quality control, hypothesis generation, and confounder identification for a wide range of genomic applications. Nonlinear approaches, despite their limitations, become increasingly useful as the size of datasets increases. UMAP, in particular, reveals a wide range of features that would not be apparent using linear maps. Given its ease of use, broad applicability, and low computational cost, we propose that UMAP should become a default companion to PCA in large genomic cohorts.
